### 2021-11-11 chapter2/chapter.tex
move aD initial condition to appendix, simplify presentation.

We use the notation \(\bs a_{\ell,i}(x_0) = \bs k(x_0-y_\ell)\) for the initial orbit position corresponding to \(x_0\) when the initial phase is \(i\in\calS_+\) and \(x_0\in\calD_{\ell,i}\). Similarly, for \(i\in\calS_-\) and \(x_0\in\calD_{\ell,i}\) define the notation \(\bs a_{\ell,i}(x_0) = \bs k(y_{\ell+1}-x_0)\).

@@

>>>>

We use the notation \(\bs a_{\ell,i}(x_0) = \bs a_{\ell,+}(x_0) = \bs k(x_0-y_\ell)\) for the initial orbit position corresponding to \(x_0\) when the initial phase is \(i\in\calS_+\) and \(x_0\in\calD_{\ell,i}\). Similarly, for \(i\in\calS_-\) and \(x_0\in\calD_{\ell,i}\) define the notation \(\bs a_{\ell,i}(x_0) = \bs a_{\ell,-}(x_0) = \bs k(y_{\ell+1}-x_0)\).

@@

\paragraph{Initial conditions for phases \(i\in\calS_0\)}
While the augmented state-space scheme described above is a convergent one, the analysis of the QBD-RAP scheme is \emph{greatly} simplified if we allow initial conditions which allocate \(0\) probability to phases in \(i\in\calS_{+0}\cup\calS_{-0}\) only. To see why initial mass in \(\mathcal S_{+0}\cup\calS_{-0}\) might introduce further complexity, consider a sample path with \(\varphi(0)=k\in\calS_{+0}\) and \(\bs A(0)=\bs a\in\mathcal A\). On \(\varphi(u)\in\calS_{+0},\) for all \(u\in[0,t)\), then \(\bs A(u) = \bs a \) is constant. Upon the phase leaving \(\calS_{+0}\), at time \(t\), say, the phase will either jump to \(\calS_+\) or \(\calS_-\). If \(\varphi(t)\) jumps to \(i\in\calS_+\), the orbit process at time \(t\) will be \(\bs A(t)=\bs a \) and the process evolve from time \(t\) as if it has started in phase \(i\in\calS_+\) with initial orbit \(\bs a \) -- there are no major issues in this case. If, however, \(\varphi(t)\) jumps to \(i\in\calS_-\), then the orbit process will jump to \(\bs A(t) = \bs a \bs D\) at time \(t\). From time \(t\) onward the QBD-RAP process will evolve as if it has started in phase \(i\in\calS_-\) with initial orbit \(\bs a \bs D\). With the way we have structured the proof of convergence, we then have to complete two analyses of the QBD-RAP process -- one for the initial condition \(\bs a\) and another for the initial condition \(\bs a\bs D\).

A solution to this problem is to augment a set of ephemeral states to the QBD-RAP which capture the initial sojourn in \(\calS_0\) only. Let us denote this set of ephemeral states \(\calS_0^*\) which has the same phase dynamics as \(\calS_0\). Suppose we wish to approximate a fluid queue with the initial condition \(\mu_k\) where \(\mu_k\) has mass 1. At time \(0\), we start the QBD-RAP in the ephemeral set of states \(\calS_0^*\) in phase \(k^*\in\calS_0^*\). While in \(\calS_0^*\), the phase process evolves according to the generator \(\bs T_{00}\). Upon exiting the ephemeral set \(\calS_0^*\) at time \(t\), say, the QBD-RAP process jumps to \(L(t)=\ell,\, \bs A(t)=\int_{\calD_{\ell,i}}\bs a_{\ell,i}(x)\wrt \mu_k,\,\varphi(t)=i\), where \(\ell\in\mathcal K\), \(i\in\calS_+\cup\calS_-\).

Notice that we have not mentioned anything about the level or orbit process during the sojourn in \(\calS_0^*\). This is because all information about the level and phase during this time is contained within the initial condition \(\mu_k\), and the only thing that changes is the phase.

To include the ephemeral states within the generator of the QBD-RAP we can do the following, 
\begin{align*}
	&\left[\begin{array}{c|cccc}
		\bs T_{00} & \bs B_{-1}^* &\bs B_{0}^*&\bs B_{1}^* & \hdots \\\hline
		&\check{\bs B}_{0} &  \check{\bs B}_{+1} &&\\
		& \check{\bs B}_{-1}&\check{\bs B}_{0}&\bs B_{+1}& \\
		&&\bs B_{-1} & \bs B_0&\ddots \\
		&&&\ddots&\ddots 
	\end{array}\right],
\end{align*}
where 
\begin{align*}
	\bs B_{-1}^* &= \vligne{\bs T_{0-}\mu_k(\{0\}) & \bs 0} & \bs B_{k}^* &= \vligne{\bs T_{0+}\otimes \bs a_k^+ & \bs 0 & \bs T_{0-}\otimes \bs a_k^- & \bs 0},
	%
	\\\check{\bs B}_{0} &= \left[\begin{array}{cc}\bs T_{--} & \bs T_{-0}\\\bs T_{0-} & \bs T_{00}\end{array}\right], 
	& \check{\bs B}_{+1} &= \left[\begin{array}{cccc}\bs T_{-+}\otimes \bs \alpha & \bs 0 & \bs 0 & \bs 0 \\ \bs T_{0+}\otimes \bs \alpha & \bs 0 & \bs 0 & \bs 0 \end{array}\right],
	%
	\\\check{\bs B}_{-1} &= \left[\begin{array}{cc} \bs0&\bs0\\\bs0&\bs0\\(\bs C_-\bs P_{--})\otimes \bs s & (\bs C_-\bs P_{-0})\otimes \bs s \\\bs0&\bs0\end{array}\right], &&
\end{align*}
and 
\begin{align*}
	\bs a_\ell^r = \int_{\calD_{\ell,r}} \bs a_{\ell,r}(x)\wrt \mu_k,\, \ell\in\mathcal K\setminus\{-1,K+1\},\, r\in\{+,-\}. 
\end{align*}
